### 2020-10-27 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Updated tidycensus doc

```{r Setup, include = FALSE, results = 'hide', warning = FALSE}
library(knitr)
library(tidycensus)
library(tidyverse)
library(ggplot2)

Sys.getenv("CENSUS_API_KEY")
```

```{r}
acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)

view(acs5_2018_vars)
```

```{r}
sex_age_vector <- acs5_2018_vars %>%
  select(name) %>%
  slice(5:51)
```

```{r}
MI_sex_age_acs5_2018 <- get_acs(geography = "place",
              variables = c(sex_age_vector),
              state = "MI",
              year = 2018)

view(MI_sex_age_acs5_2018)
```