### 2020-10-29 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Alison updated files

```{r Setup, include = FALSE, results = 'hide', warning = FALSE}
library(knitr)
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(ggthemes)
library(jcolors)

Sys.getenv("CENSUS_API_KEY")
```

```{r}
acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)

write_csv(acs5_2018_vars, "acs5_2018_vars.csv")
```

```{r}
sex_age_vars <- acs5_2018_vars %>%
  slice(5:51)

sex_age_vec <- sex_age_vars %>%
  select(name) %>%
  unlist()

sex_age_label <- sex_age_vars %>%
  select(label) %>%
  unlist()
```

```{r}
MI_place_sex_age_acs5_2018 <- get_acs(geography = "place",
              variables = sex_age_vec,
              place = "Manistee",
              state = "MI",
              year = 2018)

Man_city_sex_age_acs5_2018 <- MI_sex_age_acs5_2018 %>%
  slice(18425:18471) %>%
  mutate(sex_age_label = sex_age_label) %>%
  select(sex_age_label, estimate)
```

```{r}
# https://github.com/walkerke/tidycensus/issues/131

years <- lst(2013, 2018)

MI_place_mob_acs5_2013_2018 <- map_dfr(
  years,
  ~ get_acs(
    geography = "place",
    variables = c("B07001_016", "B07001_033", "B07001_049", "B07001_065", "B07001_081"),
    state = "MI",
    place = "Manistee",
    year = .x
    ),
  .id = "year"
  )
```

```{r}
Man_city_mob_acs5_2013_2018 <- MI_place_mob_acs5_2013_2018 %>%
  filter(NAME == "Manistee city, Michigan") %>%
  add_column(label = "") %>%
  mutate(label = case_when(variable == "B07001_016" ~ "Same house",
                           variable == "B07001_033" ~ "Same county",
                           variable == "B07001_049" ~ "Different county",
                           variable == "B07001_065" ~ "Different state",
                           variable == "B07001_081" ~ "Moved from abroad")) %>%
  mutate(label = factor(label, levels = c("Same house", "Same county", "Different county", "Different state", "Moved from abroad")))
```

```{r}
p1 <- ggplot(data = Man_city_mob_acs5_2013_2018, aes(x = label, y = estimate, fill = year)) + 
  geom_bar(stat = "identity", position = position_dodge())
```

```{r}
p1 + theme_fivethirtyeight() +
  scale_fill_jcolors(palette = "pal5") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "top",
        legend.title = element_blank()) +
  ggtitle("Where Manistee residents lived compared to \nthe year before\n") +
  labs(caption = "\nSource: 2013 and 2018 ACS 5-year surveys")
```

```{r}
MI_place_age_mob_acs5_2013_2018 <- map_dfr(
  years,
  ~ get_acs(
    geography = "place",
    variables = c("B07002_002", "B07002_003", "B07002_004", "B07002_005", "B07002_006"),
    state = "MI",
    place = "Manistee",
    year = .x
    ),
  .id = "year"
  )
```

```{r}
Man_city_age_mob_acs5_2013_2018 <- MI_place_age_mob_acs5_2013_2018 %>%
  filter(NAME == "Manistee city, Michigan") %>%
  add_column(label = "") %>%
  mutate(label = case_when(variable == "B07002_002" ~ "Same house",
                           variable == "B07002_003" ~ "Same county",
                           variable == "B07002_004" ~ "Different county",
                           variable == "B07002_005" ~ "Different state",
                           variable == "B07002_006" ~ "Moved from abroad")) %>%
  mutate(label = factor(label, levels = c("Same house", "Same county", "Different county", "Different state", "Moved from abroad")))
```

```{r}
p2 <- ggplot(data = Man_city_age_mob_acs5_2013_2018, aes(x = label, y = estimate, fill = year)) + 
  geom_bar(stat = "identity", position = position_dodge())
```

```{r}
p2 + theme_fivethirtyeight() +
  scale_fill_jcolors(palette = "pal5") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "top",
        legend.title = element_blank()) +
  ggtitle("Median age of Manistee residents \nby where they lived the year before\n") +
  labs(caption = "\nSource: 2013 and 2018 ACS 5-year surveys")
```