### 2020-11-04 EPA:Census Team/Gretchen Walkability/Walkability Data Manip.Rmd
Updated Walkability Plots

---
title: "Walkability"
author: "Gretchen Streett"
date: "10/27/2020"
output: html_document
---

```{r}
library(leaflet)
library(rgdal)
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(stringr)
library(ggplot2)
library(plotly)

#get data
walk <- readOGR(dsn = "Natl_WI_SHP/Natl_WI.shp")
mwalk <- walk[walk$SFIPS == "26", ]
#writeOGR(mwalk, ".", "mwalk", driver="ESRI Shapefile")
mwalk2 <- readOGR(dsn = "mwalk.shp")

statefips <- read.csv("~/Documents/Columbia/Practicum - Aracelis Torres/Data/Walkability/State FIPS.csv")
countyfips <- read.csv("FIPS and County Names.csv")
```

```{r}
# Walkability map (Michigan Specific) 
## Target End state: blue water, better names, search by zip code/location, round Walkability Index Score, add points/layers for parks/schools/nature
walkmap <- leaflet(mwalk) %>%
 setView(lat = 44.2444, lng = -86.3243, zoom = 6) %>%
 addProviderTiles(providers$Stamen.TonerLite) %>%
  #add bloc borders
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333', opacity=1, 
  #chloropleth shading
  fillColor = ~colorQuantile("Greens", NatWalkInd)(NatWalkInd), 
  fillOpacity = 1, 
  #add pinpoint label
  label = ~stringr::str_c(CBSA_Name, ' ',
          formatC(NatWalkInd, big.mark = ',', format='f')),
  labelOptions = labelOptions(direction = 'auto'),
  #add hover highlight
  highlightOptions = highlightOptions(
          color='#000000', weight = 3,
          bringToFront = TRUE, sendToBack = TRUE))

walkmap2 <- walkmap %>% 
  addMiniMap() %>%
  addLegend(
  "bottomleft", 
  pal = colorQuantile("Greens", mwalk$NatWalkInd, n=5),
  values = ~NatWalkInd,
  title = htmltools::HTML("Walkability Index"),
opacity = 1, 
  labFormat = function(type, cuts, p) {
n = length(cuts)
paste0(cuts[-n], " – ", cuts[-1])}
)

walkmap3 <- walkmap2 %>%
    addEasyButton(easyButton(
    icon="fa-globe", 
    title="Zoom to Entire US Map",
    onClick=JS("function(btn, map){ map.setZoom(2.5); }"))) 
walkmap3
```

```{r}
# Comparison of walkability: Manistee to State to comparable rural areas (CRA)
# EQ: Should Manistee county be removed from other categories?, What constitutes "Comparable Rural Area", # Are there areas that we don't expect to be walkable and should thus be excluded? 
# Target End state: Interactive, compare based on alt. features/demographics

#Subset Manistee county block groups (FIPS code: 026101)
manwalk <- walk[(walk$GEOID10 >= "261010000000" & walk$GEOID10 <= "261030000000"), ]
manwalk <- manwalk@data
manwalk <- manwalk %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
manwalk <- manwalk %>% mutate("Location" = "Manistee County" )

# Subset block groups with similar Population Density 
mwalkdata <- mwalk@data
mwalkdata <- mwalkdata %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
mwalkdata <- mwalkdata[(mwalkdata$`Population Density` >= .8 & mwalkdata$`Population Density` <= 1.6), ]
mwalkdata <- mwalkdata %>% mutate("Location" = "Comparable Rural Area" )

# Compare to Michigan as a Whole
mwalkdatawhole <- mwalk@data
mwalkdatawhole <- mwalkdatawhole %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
mwalkdatawhole <- mwalkdatawhole %>% mutate("Location" = "Entire State" )

# Compare Walkability of two groups 
combined <- rbind(manwalk, mwalkdata, mwalkdatawhole)
combined <- combined %>% group_by(`Location`) %>%
            summarize("Walkability" = mean(`NatWalkInd`))

myplot <- ggplot(data = combined, aes(x=`Location`, y=`Walkability`, fill=`Location`)) + 
  geom_bar(stat="identity", width=0.5) +
  scale_fill_brewer(palette="Dark2") +
  scale_x_discrete(limits=c("Manistee County", "Comparable Rural Area", "Entire State")) +
  geom_text(aes(label= round(`Walkability`, 2)), vjust=1.6, color="white", size=3.5)+
  labs(title="Walkability by Location")+
  theme_minimal() + 
  theme(legend.position="none")
myplot

```

```{r}
# Country Comparison

# Join State Names to Data 
allwalk <- walk@data
allwalk <- separate(allwalk, GEOID10, into = "State Code", sep= 2, remove = FALSE)
allwalk$`State Code` <- as.numeric(allwalk$`State Code`)
walkwithstates <- full_join(allwalk, statefips, by = c("State Code" = "FIPS"))
walkwithstates <- walkwithstates %>% filter(`State Code` <= 56)
walkwithstates <- walkwithstates %>% drop_na(`State Code`)
walkwithstates <- walkwithstates %>% drop_na(`State`)

# Summarize for plot
walkwithstates <- walkwithstates %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
walkwithstates <- walkwithstates %>% 
  mutate(
      CRA = case_when(
      (`Population Density` >= .8 & `Population Density` <= 1.6) ~ "Comparable Rural Area",
      TRUE ~ "Not Comparable"))
walkfullstates<- walkwithstates %>% group_by(`State`) %>% 
              summarize("Walkability" = mean(`NatWalkInd`))
walkstates<- walkwithstates %>% group_by(`State`, `CRA`) %>% 
              summarize("Walkability" = mean(`NatWalkInd`))

# Plot

#Reorder Walkability in table to make chart pretty
walkfullstates<- transform(walkfullstates, `State` = reorder(`State`, -Walkability))
#plot it 
 plota <- ggplot(walkfullstates, aes(x=State, y=Walkability)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  coord_flip() +
  labs(title="Walkability by State",
       caption="source: EPA") 
plota

#Reorder Walkability for CRA 
walkstates <- walkstates %>% filter(CRA == "Comparable Rural Area")
walkstates<- transform(walkstates, `State` = reorder(`State`, -Walkability))
#plot it 
plota.2 <- ggplot(walkstates, aes(x=State, y=Walkability)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  coord_flip() +
  labs(title="Comparable Rural Area Walkability by State",
       caption="source: EPA") 
plota.2

# plot Walkability by County
# just Michigan 
walkmi <- walkwithstates %>% filter(`State Code` == 26)
walkmi$countyfips <- paste(walkmi$`State Code`, walkmi$CFIPS, sep = "")
countyfips$X1001 <- as.character(countyfips$X1001)
walkmi <- left_join(walkmi, countyfips, by = c("countyfips" = "X1001"))
walkmi <- walkmi %>% group_by(Autauga) %>% 
              summarize("Walkability" = mean(`NatWalkInd`))
colnames(walkmi) <- c("County", "Walkability")
#write_csv(walkmi, "walkmi.csv")

#Plot it 
#Reorder Walkability in table to make chart pretty
walkmi <- transform(walkmi, `County` = reorder(`County`, -Walkability))
#plot it 
 plotc <- ggplot(walkmi, aes(x=County, y=Walkability)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  coord_flip() +
  labs(title="Walkability by County",
       caption="source: EPA") + 
  theme(axis.text.y=element_text(size= 5))
ggplotly(plotc)

# lollipop Test: is it better? 
# plotb <- ggplot(walkfullstates, aes(x=State, y=Walkability)) + 
#   geom_point(size=3) + 
#   coord_flip() +
#   geom_segment(aes(x=State, 
#                    xend=State, 
#                    y=0, 
#                    yend=Walkability)) +
#   labs(title="Ordered Bar Chart", 
#        subtitle="Walkability by State", 
#        caption="source: EPA") +
#   theme_bw()
# plotb
```