### 2020-11-04 EPA:Census Team/Manistee Bband Access Map/Data Manipulation/Broadband Data Manip.Rmd
Add Broadband Access Map for Manistee County

---
title: "Broadband Rough Edit"
author: "Gretchen Streett"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgdal)
library(tidyverse)
library(stringr)
library(leaflet)
library(RColorBrewer)

# Get Data: MI Shapefiles, MI Broadband
mwalk2 <- readOGR(dsn = "/Users/gretchenilsestreett/Documents/Columbia/Practicum - Aracelis Torres/Data/Walkability/mwalk.shp")
broadband <- read_csv("FCC MI Broadband.csv")
broadband$BlockCode <- as.character(broadband$BlockCode)
maniblocks <- readOGR(dsn = "/Users/gretchenilsestreett/Documents/Columbia/Practicum - Aracelis Torres/Data/MI Broadband/Manistee Census Blocks/Manistee County Blocks/tl_2010_26101_tabblock10.shp")

# get # of Providers per block 
mani_bband <- left_join(maniblocks@data, broadband, by = c("GEOID10" = "BlockCode"))
mani_bband <- mani_bband[, c("GEOID10", "ProviderName")]
mani_bband <- mani_bband %>% group_by(GEOID10) %>% summarize("ProvideNum" = n())
maniblocks@data <- full_join(mani_bband, maniblocks@data, by = c("GEOID10" = "GEOID10"))
#writeOGR(maniblocks, ".", "manistee bband blocks", driver="ESRI Shapefile")

#plot
#define bins 
bins <- c(1:9)
pal <- colorNumeric("YlGnBu", domain = maniblocks@data$ProvideNum)

bbandmap <- leaflet(maniblocks) %>%
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333',  
  fillColor = ~pal(ProvideNum), 
  fillOpacity = 1,
  #add pinpoint label
  label = ~stringr::str_c(NAME10, ':',
          formatC(ProvideNum)),
  labelOptions = labelOptions(direction = 'auto'),
  #add hover highlight
  highlightOptions = highlightOptions(
          color='#000000', weight = 3,
          bringToFront = TRUE, sendToBack = TRUE))

bbandmap2 <- bbandmap %>% 
  addLegend("bottomleft", 
  pal = pal, 
  values = ~ProvideNum, 
  title = htmltools::HTML("Number of Providers"),
  setView(bbandmap2, lng = -86.3242, lat = 44.2444, zoom = 14))
bbandmap2
```