### 2020-11-09 EPA_Census Team/Alison_Sociodemographic_Visualizations/_Old/ACS_5-Year_2014-2018_Social_Demographic_Characteristics_Manistee.Rmd

```{r Setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
library(tidyverse)
library(ggplot2)
```

```{r}
acs_social <- read_csv("ACS_5-Year_2014-2018_Social_Characteristics_Manistee.csv")
acs_demo <- read_csv("ACS_5-Year_2014-2018_Demographic_Characteristics_Manistee.csv")
```

```{r}
acs_social_clean <- acs_social %>%
  rename(Michigan = `Michigan!!Estimate`) %>%
  rename(Manistee_City = `Manistee city, Michigan!!Estimate`) %>%
  select(Label, Michigan, Manistee_City)

acs_demo_clean <- acs_demo %>%
  rename(Michigan = `Michigan!!Estimate`) %>%
  rename(Manistee_City = `Manistee city, Michigan!!Estimate`) %>%
  select(Label, Michigan, Manistee_City)
```

```{r}
acs_social_clean_wide <- acs_social_clean %>%
  select(Label, Manistee_City) %>%
  slice(2:19) %>%
  spread(key = Label, value = Manistee_City)
```

```{r}
acs_social <- read_csv("ACS_5-Year_2014-2018_Social_Characteristics_Manistee_WIDE.csv")
acs_demo <- read_csv("ACS_5-Year_2014-2018_Demographic_Characteristics_Manistee_WIDE.csv")
```

```{r}
acs_demo_age <- acs_demo %>%
  slice(7) %>%
  select(`SEX AND AGE!!Total population!!Under 5 years`:`SEX AND AGE!!Total population!!85 years and over`)

acs_demo_age_clean <- acs_demo_age %>%
  rename(`Under 5` = `SEX AND AGE!!Total population!!Under 5 years`) %>%
  rename(`5 to 9` = `SEX AND AGE!!Total population!!5 to 9 years`) %>%
  rename(`10 t0 14` = `SEX AND AGE!!Total population!!10 to 14 years`) %>%
  rename(`15 to 19` = `SEX AND AGE!!Total population!!15 to 19 years`) %>%
  rename(`20 to 24` = `SEX AND AGE!!Total population!!20 to 24 years`) %>%
  rename(`25 to 34` = `SEX AND AGE!!Total population!!25 to 34 years`) %>%
  rename(`35 to 44` = `SEX AND AGE!!Total population!!35 to 44 years`) %>%
  rename(`45 to 54` = `SEX AND AGE!!Total population!!45 to 54 years`) %>%
  rename(`55 to 59` = `SEX AND AGE!!Total population!!55 to 59 years`) %>%
  rename(`60 to 64` = `SEX AND AGE!!Total population!!60 to 64 years`) %>%
  rename(`65 to 74` = `SEX AND AGE!!Total population!!65 to 74 years`) %>%
  rename(`75 to 84` = `SEX AND AGE!!Total population!!75 to 84 years`) %>%
  rename(`85 and older` = `SEX AND AGE!!Total population!!85 years and over`)
```

```{r}
ggplot(acs_demo_age_clean) +
  geom_bar(aes(estimate))
```

```{r}
acs_demo_Manistee <- acs_demo %>%
  slice(7)
```