### 2020-11-09 EPA_Census Team/Getting_started_in_tidycensus.Rmd

### Getting started in tidycensus

Adapted from: https://walker-data.com/tidycensus/articles/basic-usage.html

```{r}
library(tidyverse)
library(tidycensus)
```

tidycensus uses the Census API, so you will need to request a key: https://api.census.gov/data/key_signup.html

To keep your API key private, run the "census_api_key" function:

```{r}
census_api_key("YOUR API KEY GOES HERE", overwrite = FALSE, install = TRUE)
```

From then on you can simply call the key from your system environment:

```{r}
Sys.getenv("CENSUS_API_KEY")
```

To search through all available variables for a given survey and year, use the "load_variables" function:

```{r}
acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
```

You can also export the CSV in order to view the full text of each cell:

```{r}
write_csv(acs5_2018_vars, "acs5_2018_vars.csv")
```

Here is an example call for Manistee County:

```{r}
Man_county_pop_hous <- get_acs(geography = "county",
                          variables = c(population = "B00001_001", housing_units = "B00002_001"),
                          state = "Michigan",
                          county = "Manistee",
                          survey = "acs5",
                          year = 2018)
```

Note that the FIPS code for Michigan is "26" and for Manistee County is "101", so the full GEOID is "26101". You can also call these FIPS codes directly instead of the place names:

```{r}
Man_county_pop_hous <- get_acs(geography = "county",
                          variables = c(population = "B00001_001", housing_units = "B00002_001"),
                          state = "26",
                          county = "101",
                          survey = "acs5",
                          year = 2018)
```