### 2020-11-11 KPMG Team/Dashboard/visualization_scripts/cities.Rmd
cities area

---
title: "Cities (Deaths & Case)"
output: html_notebook
author: Louisa Ong (QMSS Team)
---

```{r}
```{r}
# ----------- LOAD PACKAGES ------------- #

library(tidyverse)
library(tidycensus) # US census bureau data
library(zoo) # rolling mean

# plotting 
library(ggplot2)
# library(hrbrthemes) 
# library(geofacet)

# ----------- LOADING DATASETS ----------- #

#nytimes county-level cases & deaths data from online link 
nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")

# don't need it, unless we want more variables
# census data API, variables can be changed here 
# census <- get_acs(geography = "county",
#                    variables = c(population = "B01003_001E"), # total population), # black veterans 
#                    output = "wide",
#                    county = "",
#                    year = 2018)

# US City & FIPS code boundaries 
# https://simplemaps.com/data/us-cities license 
city <- read_csv("uscities.csv")

View(city)
names(city)
```

```{r}
names(city)

# ------------------------------------------------------- # 
# county covid sum cases & deaths 
county_sum <- nyt %>%
  filter(date == max(nyt$date))

# ---------------------------------------------- #
# -----------  DATA MANIPULATION --------------- # 
# ---------------------------------------------- #

# CITY DATASET 
data_city <- city %>% 
  rename(fips = county_fips) %>% 
  group_by(city) %>% 
  mutate(city_population = sum(population))

View(nyt)

# CITY & COVID MERGE 
# only need county to merge both datasets, will keep just city areas 
city_covid_merge <- data_city %>% 
  left_join(nyt, by = "fips") %>% 
  # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date(Sys.Date() - 8))

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
city_covid_cases <- data.frame(city_covid_merge) %>% 
  group_by(city, date) %>% 
  dplyr::mutate(cases_sum = max(cases)) %>% # choose max cases for each date because some dates report more than once
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

View(city_covid_cases)
# same as above, but for deaths
msa_covid_deaths <- data.frame(msa_covid_merge) %>% 
  group_by(msa, date) %>% 
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
# contains
msa_covid <- msa_covid_merge %>% 
  left_join(msa_covid_cases, by = c("msa", "date")) %>% 
  left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
  arrange(desc(msa_population))

```