### 2020-11-12 EPA_Census Team/Bband Tables Simp/Broadband Tables.Rmd
Add Simplified maps and Bband Tables

---
title: "Broadband Tables"
author: "Gretchen Streett"
date: "11/12/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(DT)

#Import Residential Broadband Data 
MI_bband_res <- read.csv("MI_bband_res.csv")

#Import Commercial Broadband Data
MI_bband_com <- read.csv("MI_bband_com.csv")
```

```{r Residential Broadband Table}
#Residential Table Function
databycounty <- function(input){
df_display <- MI_bband_res %>% filter(County == input)
df_display <- df_display %>% group_by(ProviderName, `Technology.of.Transmission`) %>% summarize("Blks" = n(), "MaxDown" = max(MaxAdDown), "MaxUp" = max(MaxAdUp))
df_display <- df_display %>% arrange(desc(Blks))
colnames(df_display) <- c("Provider Name", "Technology", "Blocks", "MaxDown", "MaxUp")
datatable(df_display)
}

#Call function (takes County name as character string, i.e. "Manistee")
databycounty("Manistee")
```

```{r Commercial Broadband Table}
# Commercial Table Function 
bizdatabycounty <- function(input){
df_display <- MI_bband_com %>% filter(County == input)
df_display <- df_display %>% group_by(ProviderName, `Technology.of.Transmission`) %>% summarize("Blks" = n(), "MaxDown" = max(MaxCIRDown), "MaxUp" = max(MaxCIRUp))
df_display <- df_display %>% arrange(desc(Blks))
colnames(df_display) <- c("Provider Name", "Technology", "Blocks", "MaxDown", "MaxUp")
datatable(df_display)
}

#Call function
bizdatabycounty("Benzie")
```