### 2020-11-12 EPA_Census Team/Simplified Maps/Simplified Maps.Rmd

---
title: "Simplified Maps"
author: "Gretchen Streett"
date: "11/12/2020"
output: html_document
---

``` {r}
library(tidyverse)
library(tmap)
library(ggplot2)
library(sf)
library(ggthemes)
library(viridis)

MIdata <- read.csv("MIbbandwalkdata")
```

```{R Walkability Map}
#Function
walkbycounty <- function(input){
subsetdata <- MIdata %>% filter(V2 == input)
ggplot() + geom_polygon(data = subsetdata, aes(x=long, y=lat, group=group, fill=NatWalkInd), color = "black") +
  scale_fill_gradientn(colours=c("royalblue", "green"), 
           name="Walkability") + theme_map()
}

#call function (Takes County Name as Character String, i.e. "Manistee")
walkbycounty("Emmet")
```

```{R Broadband Map}
# Mapping Function:
bbandbycounty <- function(input){
subsetdata <- MIdata %>% filter(V2 == input)
ggplot() + geom_polygon(data = subsetdata, aes(x=long, y=lat, group=group, fill=GroupProvideNum), color = "black") +
  scale_fill_viridis(option = "A", name="Number of \n Broadband Providers") + theme_map()
}

#Call Function (Takes County Name as a character string, I.e. "Manistee")
bbandbycounty("Manistee")
```