### 2020-11-24 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd

---
title: "Visualizations: Metropolitan Areas per Capita, Wave I, II, III (Deaths)"
output: html_notebook 
---

# CONTENTS OF THIS SCRIPT

### Loading Packages 
### Datasets 
### Merging County data to Top 50 populated MSA & Region Dataset (for spatial attributes)
### Adding Spatial Attributes 
### Filtering by waves
### Visualization in Maps

```{r message = FALSE}
# ----------- LOAD PACKAGES ------------- #

library(tidyverse)
library(tidycensus)

# plotting 
library(ggplot2)
library(ggthemes)
library(showtext)

showtext_auto()

font_add(family = "raleway", regular = "C:/Users/louis/Dropbox/Stuff/Fonts/Raleway-Regular.ttf")

```

```{r cache = TRUE}

# ----------- LOADING DATASETS ----------- #

# MSA Things
# from Sydney
msa_index <- read_csv("data/msa/msa_mix_index.csv") # csa and msa with population 
covid <- read_csv("data/msa/us_daily_msa_split_2020-11-16.csv") # split

# metropolitan state areas delineation from file taken off Census Bureau
msa <- read_csv("data/msa.csv") # using this bc FIPS code needs to be 5 digits

# -------------- MSA_INDEX dataset check
# top 50 most populated csa and msas
# msa_index %>% 
#   select("mix_index") %>% 
#   distinct() %>% 
#   count()

# -------------- MSA_SPLIT dataset check
# Regional minus State count 
# CSA/MSA count

# View(msa)
```

# Adding County Data (to match with spatial lines)

```{r}

# need to merge county back in to get the lines

# # rename to get 5-digit fips, select only needed variables, remove NAs
msa_w_fips <- msa %>%
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>% #
  select(fips, `CSA Title`, `County/County Equivalent`, `State Name`)

# # left join index of 50 with fips codes
# msa_index_fips <- msa_index %>% 
#   left_join(msa_w_fips, by = c("CSA Title", "County/County Equivalent", "State Name")) %>% 
#   rename(county = `County/County Equivalent`) %>% 
#   rename(csa = `CSA Title`) %>%
#   rename(state = `State Name`) %>%
#   rename(agg_index = mix_index)

msa_index_2 <- msa_index %>% 
  rename(county = `County/County Equivalent`) %>% 
  rename(csa = `CSA Title`) %>%
  rename(state = `State Name`) %>%
  rename(agg_index = mix_index) %>% 
  rename(fips = FIPS_combined) %>% 
  names()

# anti join to get the counties not in msa_index 
region_fips <- msa_w_fips %>% 
  anti_join(msa_index, by = "CSA Title") %>% 
  rename(county = `County/County Equivalent`) %>%
  mutate(agg_index = `State Name`) %>%
  rename(state = `State Name`) %>%
  rename(csa = `CSA Title`)

# joining fips info with covid dataset (has states + selected msas)
state_msa_fips_join <- covid %>% 
  full_join(msa_index_fips, by = "agg_index")# %>%

# keeping only msas and counties, since they have fips
msa_covid_only <- state_msa_fips_join %>%
  drop_na(fips)

# keeping only states, don't have counties
state_covid_only <- state_msa_fips_join %>% 
  anti_join(msa_covid_only) %>% 
  select(-fips, -mix_index_population, -FIPS_population, -FIPS_combined, -`Central/Outlying County`, -state, -county, -csa, -`Metropolitan Division Title`)

names(msa_region)

region_fips_join <- msa_region %>% 
  full_join(region_fips, by = "agg_index") %>%
  drop_na(fips)

View(msa_covid_only)

View(region_fips_join)
View(covid)

final_daily <- msa_fips_join %>% # or final total 
  bind_rows(region_fips_join) %>% 
  select(index, Date, state, agg_index, fips, csa, county, population, 
         Confirmed, Deaths, new_confirmed, new_deaths, 
         new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
         per_100k_new_confirmed_7, per_100k_new_deaths_7)

names(final_daily)
# write.csv(final_daily, "data/final-50-county-daily.csv")

```

# Adding spatial attributes

```{r cache = TRUE}

# Adding spatial attributes

library(sf)
library(jsonlite) # from FIPS -> Geo (lat, long)
library(gridExtra) # for grid plot

county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")

# merge county 
# add FIPS as a variable from State Fips and County Fips
county_lines <- county_lines %>% 
  mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))

# add lat long attributes to county dataset          
county_geo <- merge(county_lines, final_daily, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

map_daily <- county_map # daily covid map

```

# Wave/Dates Filter  
### For total cases

```{r}
# library(spdplyr) # spatial version of dplyr

# Dates of Waves

wave_2 <- as.Date("2020-06-15") # start of wave 2
wave_3 <- as.Date("2020-10-15") # start of wave 3

# all cases map 
map_total <- county_map %>% 
  filter(Date == max(as.Date(county_map$Date), na.rm = TRUE)) %>% # need to remove NA
  rename(state_fips = STATEFP) %>%
  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)

# wave 2 start date 
map_w2_start <- map_daily %>%
  filter(Date == "2020-06-15") %>% 
  rename(state_fips = STATEFP) %>%
  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)

# wave 2 map 
map_w2 <- map_total %>% 
  st_join(map_w2_start, by = "fips") %>% 
  mutate(Confirmed = Confirmed.x - Confirmed.y) %>% 
  mutate(Deaths = Deaths.x - Deaths.y) %>% 
  rename(start_date = Date.y) %>% 
  rename(end_date = Date.x) %>% 
  rename(fips = fips.x) %>% 
  rename(county = NAME.x) %>% 
  rename(population = population.x) %>% 
  rename(csa = csa.x) %>% 
  rename(agg_index = agg_index.x) %>% 
  rename(state = state.x) %>%
  rename(index = index.x) %>%
  rename(county_fips = COUNTYFP.x) %>%
  rename(state_fips = state_fips.x) %>%
  select(fips, index, agg_index, state, county, csa, population, Confirmed, Deaths, start_date, end_date, state_fips, county_fips, geometry, LSAD.x, ALAND.x, AWATER.x)

```

# Visualizations (finally)

```{r}
library(usmap) # to get US Map Data

names(map_w2)
fips("New York")
#state_name,

# county state map 
mapping <- function(map_type, pax) {
  # 
  # fips_code <- fips(state = state_name) 
  # 
  # map <- map_type %>% 
  #   filter(state_fips == fips_code) %>% 
  #   filter(Deaths >= 0)
  #

  map <- map_type

  ggplot(data = map) +
    geom_sf(aes(fill = Deaths, color = Deaths)) + 
    scale_fill_viridis_c(option = "viridis", direction = -1) + 
    labs(title = "something", #paste(state),
         subtitle = "Total Deaths") +
    theme_void() + 
    theme(text = element_text(family = "raleway"))

  # if (pax = FALSE) {
  #   
  #   map <- 
  #   
  # } else if (pax = TRUE) {
  #   
  #   
  #

}

dim(map_w2)

length(map_total$Deaths)
length(map_w2$Deaths)

mapping(map_w2)

```