### 2020-11-25 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd
msa and region waves I, II, III

---
title: "Visualizations: Metropolitan Areas per Capita, Wave I, II, III (Deaths)"
output: html_notebook 
---

# CONTENTS OF THIS SCRIPT

### Loading Packages 
### Datasets 
### Merging County data to Top 50 populated MSA & Region Dataset (for spatial attributes)
### Adding Spatial Attributes 
### Filtering by waves
### Visualization in Maps

```{r message = FALSE}
# ----------- LOAD PACKAGES ------------- #

library(tidyverse)
library(tidycensus)

# plotting 
library(ggplot2)
library(ggthemes)
library(showtext)

showtext_auto()

font_add(family = "raleway", regular = "C:/Users/louis/Dropbox/Stuff/Fonts/Raleway-Regular.ttf")

```

```{r cache = TRUE}

# ----------- LOADING DATASETS ----------- #

# MSA Things
# from Sydney
msa_index <- read_csv("data/msa/msa_mix_index.csv") # csa and msa with population 
covid <- read_csv("data/msa/us_daily_msa_split_2020-11-16.csv") # split 
covid_state <- read_csv("data/us_state_level_clean_2020-11-20.csv") # states only

# metropolitan state areas delineation from file taken off Census Bureau
msa <- read_csv("data/msa.csv") # using this bc FIPS code needs to be 5 digits

# -------------- MSA_INDEX dataset check
# top 50 most populated csa and msas
# msa_index %>% 
#   select("mix_index") %>% 
#   distinct() %>% 
#   count()

# -------------- MSA_SPLIT dataset check
# Regional minus State count 
# CSA/MSA count

# View(msa)

covid_state
```

# Adding County Data (to match with spatial lines)

```{r}

# need to merge county back in to get the lines

# # rename to get 5-digit fips, select only needed variables, remove NAs
msa_w_fips <- msa %>%
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>% #
  select(fips, `CSA Title`, `County/County Equivalent`, `State Name`)

msa_index_2 <- msa_index %>% 
  rename(county = `County/County Equivalent`) %>% 
  rename(csa = `CSA Title`) %>%
  rename(state = `State Name`) %>%
  rename(agg_index = mix_index) %>% 
  rename(fips = FIPS_combined)

# joining fips info with covid dataset (has states + selected msas)
state_msa_fips_join <- covid %>% 
  full_join(msa_index_2, by = "agg_index")# %>%

# keeping only msas and counties, since they have fips
msa_covid_only <- state_msa_fips_join %>%
  drop_na(fips)

# keeping only states, don't have counties
state_covid_only <- state_msa_fips_join %>% 
  anti_join(msa_covid_only) %>%
  select(-fips, -mix_index_population, -FIPS_population, -`Central/Outlying County`, -state, -county, -csa, -`Metropolitan Division Title`)

```

# Getting Wave Count

## for States

```{r}
wave_2 <- as.Date("2020-06-15") # start of wave 2
wave_3 <- as.Date("2020-10-15") # start of wave 3

state_covid_reg <- covid_state %>% 
  rename("NAME" = "Province_State") %>%
  filter(! NAME %in% c("Alaska", "Hawaii", "Rhode Island", "District of Columbia"))

state_reg_total <- state_covid_reg %>% 
  filter(Date == max(as.Date(state_covid_reg$Date), na.rm = TRUE)) %>% # need to remove NA
  mutate(Date_total = Date)

state_reg_w2_start <- state_covid_reg %>% 
  filter(Date == wave_2) %>%
  mutate(Date_W2 = Date)

state_reg_w3_start <- state_covid_reg %>% 
  filter(Date == wave_3) %>% 
  mutate(Date_W3 = Date)

View(state_reg_w2_start)
View(state_reg_total)
View(state_all)

state_reg_all <- state_reg_total %>% 
  rename(Confirmed_total = Confirmed) %>% # Start - Latest
  rename(Deaths_total = Deaths) %>%
  # joining wave 2 and beyond numbers
  left_join(state_reg_w2_start, by = "NAME") %>%  # Start - June 15
  # wave 1 only %>% 
  rename(Confirmed_W1 = Confirmed) %>%  # Start - June 15
  rename(Deaths_W1 = Deaths) %>% # Start - June 15
  # wave 2 and beyond column 
  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed_W1) %>% # June 15 - Latest
  mutate(Deaths_W2_plus = Deaths_total - Deaths_W1) %>% # June 15 - Latest
  # joining wave 3 numbers 
  left_join(state_reg_w3_start, by = "NAME") %>% # Start - Oct 15 ie Wave 1 & Wave 2
  # only wave 2 
  mutate(Deaths_W2_only = Deaths - Deaths_W1) %>% # June 15 - Oct 15
  mutate(Confirmed_W2_only = Confirmed - Confirmed_W1) %>% # June 15 - Oct 15
  # wave 3 and beyond
  mutate(Deaths_W3 = Deaths_total - Deaths) %>%  # Oct 15 - Latest
  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>% # Oct 15 - Latest
  # tidy up, selecting only relevant variables
  select(NAME, population, 
         Date_total, Confirmed_total, Deaths_total, 
         Confirmed_W1, Deaths_W1,
         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only, 
         Date_W3, Confirmed_W3, Deaths_W3, stringency, containment, RegionCode)

```

## for States without MSA

```{r}

state_covid_only_2 <- state_covid_only %>% 
  rename("NAME" = "agg_index") %>%
  filter(! NAME %in% c("Alaska", "Hawaii", "Rhode Island", "District of Columbia"))

state_total <- state_covid_only_2 %>% 
  filter(Date == max(as.Date(state_covid_only_2$Date), na.rm = TRUE)) %>% # need to remove NA
  mutate(Date_total = Date)

state_w2_start <- state_covid_only_2 %>% 
  filter(Date == max(wave_2)) %>%
  mutate(Date_W2 = Date)

state_w3_start <- state_covid_only_2 %>% 
  filter(Date == max(wave_3)) %>% 
  mutate(Date_W3 = Date)

state_all <- state_total %>% 
  rename(Confirmed_total = Confirmed) %>% # Start to Latest
  rename(Deaths_total = Deaths) %>%
  # joining wave 2 and beyond numbers
  left_join(state_w2_start, by = "NAME") %>%  # Start to June 15
  # wave 1 only %>% 
  rename(Confirmed_W1 = Confirmed) %>%  # Start to June 15
  rename(Deaths_W1 = Deaths) %>% # Start to June 15
  # wave 2 and beyond column 
  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed_W1) %>% # June 15 to Latest
  mutate(Deaths_W2_plus = Deaths_total - Deaths_W1) %>% # June 15 to Latest
  # joining wave 3 numbers 
  left_join(state_w3_start, by = "NAME") %>% # Start to Oct 15 ie Wave 1 & Wave 2
  # only wave 2 
  mutate(Confirmed_W2_only = Confirmed - Confirmed_W1) %>% # June 15 to Oct 15
  mutate(Deaths_W2_only = Deaths - Deaths_W1) %>% # June 15 to Oct 15
  # wave 3 and beyond
  mutate(Deaths_W3 = Deaths_total - Deaths) %>%  # Oct 15 to Latest
  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>% # Oct 15 to Latest
  # tidy up, selecting only relevant variables
  select(index, NAME, population, 
         Date_total, Confirmed_total, Deaths_total, 
         Confirmed_W1, Deaths_W1,
         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only, 
         Date_W3, Confirmed_W3, Deaths_W3)

View(msa_all)

```

## for MSAs/Top 50

```{r}

msa_covid_only_2 <- msa_covid_only %>% 
  rename("NAME" = "agg_index")

msa_total <- msa_covid_only_2 %>% 
  filter(Date == max(as.Date(msa_covid_only_2$Date), na.rm = TRUE)) %>% # need to remove NA
  mutate(Date_total = Date)

msa_w2_start <- msa_covid_only_2 %>% 
  filter(Date == max(wave_2)) %>%
  mutate(Date_W2 = Date)

msa_w3_start <- msa_covid_only_2 %>% 
  filter(Date == max(wave_3)) %>% 
  mutate(Date_W3 = Date)

msa_all <- msa_total %>% 
  rename(Confirmed_total = Confirmed) %>% # Start - Latest
  rename(Deaths_total = Deaths) %>%
  # joining wave 2 and beyond numbers
  left_join(msa_w2_start, by = "NAME") %>%  # Start - June 15
  # wave 1 only %>% 
  rename(Confirmed_W1 = Confirmed) %>%  # Start - June 15
  rename(Deaths_W1 = Deaths) %>% # Start - June 15
  # wave 2 and beyond column 
  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed_W1) %>% # June 15 - Latest
  mutate(Deaths_W2_plus = Deaths_total - Deaths_W1) %>% # June 15 - Latest
 left_join(msa_w3_start, by = "NAME") %>% # Start - Oct 15 ie Wave 1 & Wave 2
  # only wave 2 
  mutate(Deaths_W2_only = Deaths - Deaths_W1) %>% # June 15 - Oct 15
  mutate(Confirmed_W2_only = Confirmed - Confirmed_W1) %>% # June 15 - Oct 15
  # wave 3 and beyond
  mutate(Deaths_W3 = Deaths_total - Deaths) %>%  # Oct 15 - Latest
  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>% # Oct 15 - Latest
  # tidy up, rename
  rename(metro_title = "Metropolitan Division Title") %>%
  rename(central_outlying = "Central/Outlying County") %>%
  # selecting only relevant variables
  select(index, NAME, fips, csa, county, metro_title, population, FIPS_population, mix_index_population,
         Date_total, Confirmed_total, Deaths_total,
         Confirmed_W1, Deaths_W1,
         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only,
         Date_W3, Confirmed_W3, Deaths_W3) %>%
  distinct()

```

# Adding spatial attributes

```{r cache = TRUE}

# Adding spatial attributes

library(sf)
library(jsonlite) # from FIPS -> Geo (lat, long)
library(gridExtra) # for grid plot

county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")
state_lines <- st_read("data/state_shape/cb_2018_us_state_500k.shp")

# merge county 
# add FIPS as a variable from State Fips and County Fips
county_lines <- county_lines %>% 
  mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))

# add lat long attributes to county dataset          
county_geo <- merge(county_lines, msa_all, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

# merge state 
# add lat long attributes to state dataset

state_geo <- merge(state_lines, state_all, by = "NAME")
state_map <- state_geo %>% st_transform('+proj=longlat +datum=WGS84')

state_reg_geo <- merge(state_lines, state_reg_all, by = "NAME")
state_reg_map <- state_reg_geo %>% st_transform('+proj=longlat +datum=WGS84')

```

# Visualizations (finally)

## state

```{r}

options(scipen = 999)

library(usmap) # to get US Map Data
library(cowplot) # for grid plot

# fips("New York")
#state_name,

g_total <- ggplot() +
  geom_sf(data = state_map, 
          aes(fill = Deaths_total/population), 
          color = NA) +  # color = NA 
  geom_sf(data = county_map, aes(fill = Deaths_total/population), color = NA) + # MSA Things
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
  labs(title = "Total Deaths\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 50))

max_W1 <- max(c(state_map$Deaths_W1 / state_map$population, county_map$Deaths_W1 / county_map$population))
max_W2 <- max(c(state_map$Deaths_W2_only / state_map$population, county_map$Deaths_W2_only / county_map$population))
max_W3 <- max(c(state_map$Deaths_W3 / state_map$population, county_map$Deaths_W3 / county_map$population))

min_W1 <- min(c(state_map$Deaths_W1 / state_map$population, county_map$Deaths_W1 / county_map$population))
min_W2 <- min(c(state_map$Deaths_W2_only / state_map$population, county_map$Deaths_W2_only / county_map$population))
min_W3 <- min(c(state_map$Deaths_W3 / state_map$population, county_map$Deaths_W3 / county_map$population))

max_W1 
max_W2
max_W3
min_W1
min_W2
min_W3

max(state_map$Deaths_W1 / state_map$population)
max(county_map$Deaths_W1 / county_map$population)
min(state_map$Deaths_W1 / state_map$population)
min(county_map$Deaths_W1 / county_map$population)

max(state_map$Deaths_W2_only / state_map$population)
max(county_map$Deaths_W2_only / county_map$population)
min(state_map$Deaths_W2_only / state_map$population)
min(county_map$Deaths_W2_only / county_map$population)

max(state_map$Deaths_W3 / state_map$population)
max(county_map$Deaths_W3 / county_map$population)
min(state_map$Deaths_W3 / state_map$population)
min(county_map$Deaths_W3 / county_map$population)

g_w1 <- ggplot(data = state_reg_map) +
  geom_sf(aes(fill = Deaths_W1/population), color = NA) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W1, max_W1)) + 
  labs(title = "U.S. States - Wave 1 Deaths per Capita",
      subtitle = "Until 15 June 2020\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 30))

g_w2 <- ggplot(data = state_reg_map) +
  geom_sf(aes(fill = Deaths_W2_only/population), color = NA) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W2, max_W2)) + 
  labs(title = "U.S. States - Wave 2 Deaths per Capita",
       subtitle = "From 15 June 2020 - 15 October 2020\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 30))

g_w3 <- ggplot(data = state_reg_map) +
  geom_sf(aes(fill = Deaths_W3/population), color = NA) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W3, round(max_W3))) + 
  labs(title = "U.S. States - Wave 3 Onwards Deaths per Capita",
       subtitle = "After 15 October 2020\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 30))

# state and county msa 
g_c_w1 <- ggplot() +
  geom_sf(data = state_map, aes(fill = Deaths_W1/population), color = NA) +
  geom_sf(data = county_map, aes(fill = Deaths_W1/population), color = NA) + # MSA Things
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W1, max_W1)) + 
  labs(title = "Metropolitan State Areas, Regional Areas - Wave 1 Deaths per Capita", 
       subtitle = "Until 15 June 2020\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 30))

g_c_w2 <- ggplot(data = state_map) +
  geom_sf(aes(fill = Deaths_W2_only/population), color = NA) + # size = 0.3
  geom_sf(data = county_map, aes(fill = Deaths_W2_only/population), color = NA) + # MSA Things
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W2, max_W2)) +
  labs(title = "Metropolitan State Areas, Regional Areas - Wave 2 Deaths per Capita",
       subtitle = "From 15 June 2020 - 15 October 2020\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 30))

g_c_w3 <- ggplot(data = state_map) +
  geom_sf(aes(fill = Deaths_W3/population), color = NA) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W3, round(max_W3))) + 
    geom_sf(data = county_map, aes(fill = Deaths_W3/population), color = NA) + # MSA Things
  labs(title = "Metropolitan State Areas, Regional Areas - Wave 3 Deaths per Capita",
       subtitle = "After 15 October 2020\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 30))

# Grid Plots 
p <- plot_grid(g_total, g_w2, g_w3, ncol = 3)

p_2 <- plot_grid(g_w2, g_c_w2, ncol = 2)
p_3 <- plot_grid(g_w3, g_c_w3, ncol = 2)

title <- ggdraw() + 
  draw_label("Total Deaths per Capita in U.S. by State per Wave", fontface='bold') + 
  theme(text = element_text(family = "raleway"))

plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))

# ggsave("viz/state-waves.png", width = 8, height = 3, dpi = 300)
# ggsave("viz/state-waves-total.png", g_total, width = 8, height = 5, dpi = 300)
# ggsave("viz/state-waves-W1.png", g_w1, width = 8, height = 5, dpi = 300)
# ggsave("viz/state-waves-W2.png", g_w2, width = 8, height = 5, dpi = 300)
# ggsave("viz/state-waves-W3.png", g_w3, width = 8, height = 5, dpi = 300)

ggsave("viz/state-wave-w1.png", g_w1, width = 8, height = 5, dpi = 300)
ggsave("viz/state-msa-wave-w1.png", g_c_w1, width = 8, height = 5, dpi = 300)

ggsave("viz/state-wave-w2.png", g_w2, width = 8, height = 5, dpi = 300)
ggsave("viz/state-msa-wave-w2.png", g_c_w2, width = 8, height = 5, dpi = 300)

ggsave("viz/state-wave-w3.png", g_w3, width = 8, height = 5, dpi = 300)
ggsave("viz/state-msa-wave-w3.png", g_c_w3, width = 8, height = 5, dpi = 300)

ggsave("viz/state-msa-waves-W1.png", plot_grid(g_w1, g_c_w1, ncol = 2), width = 14, height = 5, dpi = 300)
ggsave("viz/state-msa-waves-W2.png", plot_grid(g_w2, g_c_w2, ncol = 2), width = 14, height = 5, dpi = 300)
ggsave("viz/state-msa-waves-W3.png", plot_grid(g_w3, g_c_w3, ncol = 2), width = 14, height = 5, dpi = 300)

# ggsave("viz/state-msa-waves-temp.png", width = 14, height = 5, dpi = 300)

```

## top 50

```{r}

c_total <- ggplot(data = county_map) +
  geom_sf(aes(fill = Deaths_total/population), size = 0.3) +  # color = NA 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
  labs(title = "Total Deaths\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 50))

```